Maximum Entropy Spectral Analysis: a case study

### 2021-02-24 paper/mesa.tex
Edit in validation section

\usepackage{amsmath,amssymb}
\usepackage{amsfonts}
\usepackage[normalem]{ulem}
\usepackage{textcomp}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{bm}
\usepackage{afterpage}
\usepackage{graphicx}
\usepackage{psfrag}
\usepackage{mathtools}
%\usepackage{tensor}
\usepackage{layouts}
%\usepackage{DejaVuSans}
\usepackage{epstopdf}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[utf8]{inputenc}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{tabularx}
\usepackage{ragged2e}
\usepackage{blindtext}
\usepackage{graphicx}
\usepackage{siunitx}
	\sisetup{output-decimal-marker={.}}
\newtheorem{theorem}{Theorem}
\usepackage{float}

@@

The application of MESA is not limited to spectral estimates, but it also provides a link between spectral analysis and the study
of autoregressive processes.\cite{doi:10.1029/RG013i001p00183}.
An autoregressive stationary process of order $p$ (AR($p$)) is a time series whose values satisfy the following expression: 
\begin{equation} \label{eq:AR_p}
    x_t - b_1 x_{t-1} - b_2 x_{t-2} \dots b_p x_{t - p} = \nu_t
\end{equation}
where $b_1, \ldots, b_p$ are real coefficients and $\nu_t$ is white noise with a given variance $\sigma^2$.
Thus, an AR($p$) process models the dependence of the value of the process at time $t$ on every past $p$ observations, 
thus being potentially able to model complex autocorrelation between data.
\par
To show the analogy, we compute the PSD of such process and we show that it is formally equivalent to the PSD obtained in eq.~\ref{eq:MESApsd}.
We start taking the z \sschmidt{What is the z transform} \wdp{it is a relative of the Fourier transform, z are the roots of the unit in the complex plane} transform of the previous eq.~\ref{eq:AR_p}: 
\begin{align}
    \sum_t x_t z^t - \sum_i b_i z^i\sum_t x_{t - i} z^{t - i} = \sum_t \nu_t z^t.
\end{align}
Calling $\tilde x(z)$ and $\tilde \nu (z)$, the transformed quantities, 
in the $z$ domain, the process takes the form:
\begin{equation}
    \tilde x(z) = \frac{\tilde\nu(z)}{\left(1 - \sum_{i = n}^p b_n z^n \right)}
\end{equation}
Since we assumed a stationary process, $\tilde{x}(z)$ is analytic both on and inside the unit circle. Taking its square value and evaluating it on the unit circle $z = e^{-\imath 2 \pi f \Delta t}$, from the definition of spectral density one obtains:
\begin{equation}\label{eq:ARspectrum}
    S(f) = \vert \tilde x(z)\vert ^ 2 = 
    \frac{\vert \tilde \nu(f) \vert ^ 2}{\left\vert 1 - \sum_{n = 1}^p b_i e^{\imath 2 \pi f n \Delta t} \right\vert ^ 2}.
\end{equation}
The numerator is the spectral density of white noise $\nu_t$, i.e. its (constant) variance $\sigma^2$.
Eq.~(\ref{eq:ARspectrum}) and Eq.~(\ref{eq:MESApsd}) are equivalent, if we identify $b_i = - a_i$ and $P_N \Delta t= \sigma ^ 2$.
This shows that the MAXENT estimation of the PSD computes the coefficients of the autoregressive process that is able to model the data faithfully.

@@

This section is devoted to an extensive study on how to make such choice.
First of all (\ref{sec:optimizers}), we are going to define three different ``metrics" to measure how well the algorithm is performed: we will refer to them as \textit{optimizers}. The basic idea is to monitor, as the autoregressive order considered increase, the performance of the algorithm (in some sense) and pick, among the orders the one that yields better results.
Second, the performance of different optimizers will be assessed by performing the spectral analysis on time series with given analytical PSD, sec.~\ref{sec:arp_validation}.
In a last subsection~\ref{sec:LIGO_validation} real data from LIGO-Virgo gravitational waves observatory.

\sschmidt{Qui ho provato a mettere un po' di ordine alla sezione. Ho scritto l'introduzione sopra per spiegare i passaggi logici della validation. Spero che vada bene. La parte sotto è ancora da rivedere e probabilmente c'è molto da asciugare: sono quasi 10 pagine!}

\subsection{Choosing an autoregressive order}\label{sec:optimizers} 
\sschmidt{optimizer è la parola giusta? Optimizer è un algoritmo... Si potrebbe chiamare optimizer ``metrica"?}
A good choice for the autoregressive order $M$ is proposed to be \cite{doi:10.1190/1.1440902}
\begin{equation}\label{eq:MMAx}
M_{max} = 2N / \ln{(2N)}
\end{equation}
However, this is just an upper limit for the estimate of the length $M$ and the optimal algorithm could employ much less points.
We then need a more sophisticated method for computing the right value for $M$.
Various optimizers to assess the algorithm performance have been proposed in literature. We summarize them below:

\begin{itemize}
\item \textbf{Final prediction Error (FPE)} 
The first criterion is due to Akaike (1970) \cite{Akaike1970StatisticalPI}. He proposed that M has to be chosen as the length that reduces the error when the filter is used as a predictor: 
\begin{equation}
    FPE(m) = \mathbb{E}\left[ \left((x_t - \hat x_t) ^ 2\right) \right]
\end{equation}
with $\hat{x}_t = \sum_{i = 1}^M a_i x_{t - i}$.
Minimizing FPE is the same as minimizing the quantity: 
\begin{equation}
    P_{m} \frac{N + m + 1}{N - m - 1}
\end{equation}
with $P_M$ defined by the prediction error filter equation \sschmidt{What is this equation?}.
\item \textbf{Criterion Autoregressive Transfer function (CAT)}
This second optimizer is due to Parzen \cite{bhansali1986}, and it is defined as
\begin{equation}
    CAT(m) = \frac{1}{N}\sum_{k = 1}^m \frac{N - k}{N P_k} - \frac{N - m}{N P_m}.
\end{equation}
It has the interesting property that the distance between the original filter and the estimate is minimum and asymptotically unbiased.

\item \textbf{Optimum Bayes Decision rule (OBD)} 
The last criterion we will consider is the OBD (Rao, 1982)\cite{doi:10.1029/WR018i004p01097}. With the technique of asymptotic integration, th optimizer is proposed to be the length that maximizes the posterior distribution \sschmidt{Which posterior?}: 
   \begin{align}
        OBD&(m) = (N - m - 2) \log(P_m) \nonumber\\
        &+ m \log(N) + \sum_{k = 0}^{m-1} \log(P_k) + \sum_{k = 1}^{m} a_k^2.
    \end{align}
\end{itemize}

Once an optimizer is chosen, the choice of the best recursion order is straightforward: we solve the Levinson recursion until $M_{max}$, as given in Eq.~\ref{eq:MMAx}, iterations are reached. Then, the order $M$ is selected to be the one that minimize the chosen optimizer value.

In a real implementation of the algorithm, computing all the recursion up to $M_{max}$ can result in a significant waste of computational power: the optimal value is often $M_{opt}<< M_{max}$ and in such cases, computing all the values of $m$ until $M_{max}$ is not useful.
In practice, we can apply an \textit{early stopping} procedure: every few iterations we look for the best order of $M_{opt}$; if this value does not change for a while, the computation is stopped and a good (local) minimum of the optimizer is found.

The following sections will be devoted to the study of the statistical properties of the optimizers above: we need to understand which choice provides the best quality in the reproduction of some known power spectral densities.

\subsection{Choosing an optimizer - A gaussian PSD} \label{sec:arp_validation}
We test the performance of the three optimizers on a random time series with a known power spectral density.
We take the PSD to be a Gaussian distribution with mean $\mu = 2.5 \quad units$ and standard deviation $\sigma = 0.5 \quad units$, see fig.~\ref{fig:autocorr}.
Of course, the time series will be generated by sampling a frequency vector from ${p(\tilde{n}(f_i)) \propto \exp{\left\{-\frac{1}{2}\sum_i\frac{n_i ^2}{S(f_i)}\right\}}}$ and performing the inverse Fast Fourier Transform on the frequency series.

We will apply Burg method with each optimizer on a number of simulated time series, obtaining a variety of estimate for PSD.
We will study statistical properties for both single reproductions and ensemble average. In either cases, we will quantify the agreement between the reconstructed quantities and the expected ones.
In many realistic cases, such ensemble average is not accessible (one car rarely observe multiple time series realizations for the \textit{same} process), so that single reproduction studies are probably more useful. Still, study of ensemble average is important to highlight some interesting properties. In this way we can fully characterize the performance of each optimizer .
n what follows, we will indicate with $\bar S$ any ensemble averaged quantity.

We will attentively study the role of the filter length estimate M in determining the accuracy of the reconstruction. We will consider every optimizer separately and compare result only at the end of this section.
In section~\ref{sec:LIGO_validation}, we will perform the analysis again using, for the purpose of generating the time series, the Advanced LIGO design sensitivity PSD \cite{Ligo}.

@@

\begin{figure}
    \centering
        \includegraphics[width = \linewidth]{Images/NormalPSD/NormalNoise.pdf}
    \caption{Noise in both time and frequency domain. Frequency noise behaviour is representative of the chosen spectrum, having its maximum amplitude around $f = 2.5 au$}
    \label{fig:noise}
\end{figure}

For our studies, we generate a dataset of 1000 time series of 3000 points each: see fig.~\ref{fig:noise} for an example of a time series in frequency and time domain.
For every time series we will compute the spectrum with every optimization method.
To assess the magnitude of the disagreement, we compute the relative error between the mean spectrum $\hat{S}$ and the true (analytical) spectrum $S$ for every frequency bin:
\begin{equation}
r(f) = \frac{|\hat{S}(f) - S(f)|}{S(f)}\; .
\end{equation}

We report the average values of the relative error for the different optimizers in Table~\ref{tab:relative_errors}.

\sschmidt{Sono arrivato fino a qui nelle modifiche}
\begin{table}
	\caption{Average relative error for different optimizers, computed in different ways. We consider a set of $1000$ realization of a noisy time series of $3000$ points.
	We report the average error  $\bar{r}_{\bar{S}}$ on the whole set of simulations, the error $\bar{r}_{\bar S}^{\mu \pm 3\sigma} $ averaged only on data points between $3\sigma$ and the error $\bar{r}$ on a single, randomly drawn realization.
	\sschmidt{I think we need to discuss on which performance error to include in the estimation}
	}
	\label{tab:relative_errors}
	\def\arraystretch{1.5}
	\begin{ruledtabular}
		\begin{tabular}{ l c c c }
			\multirow{2}{*}{Optimizer}& \multicolumn{3}{c}{Relative Error}\\
			& $\bar{r}_{\bar{S}}$	& $\bar{r}_{\bar S}^{\mu \pm 3\sigma} $ & $\bar{r}$\\
			\hline \hline
			FPE 			& $6.9 \; (46.9\%)$	& $7 \; (1.6\%)$ & \\ 
			\hline
			CAT 			& $4.5 \; (30.6\%)$	& $194 \; (45.3\%)$ & \\
			\hline
			OBD 				& $1.7 \; (11.6\%)$	& $206 \; (48.1\%)$ & \\
		\end{tabular}
	\end{ruledtabular}
\end{table}

\begin{figure}[H]
	\centering
	\includegraphics[width = \linewidth]{Images/NormalPSD/FPEpsdAndResiduals.pdf}
	\caption{Ensemble Average for FPE reconstruction with 90\% Credibility Regions (above) and percentage error (below)}
	\label{fig:FPEmean}
\end{figure}

\subsubsection{Final Prediction Error (FPE)}
The results of our investigation on te performance of FPE are reported in fig.~\ref{fig:FPEmean}: we plot the ``true" spectrum, the reconstructed spectrum as well as the $90\%$ credible region of the predicted spectrum.
Qualitatively, there is a good agreement between the reconstructed spectrum and the true spectrum; however, we note that the reconstruction is not very accurate in the low frequency region.

For the entire spectrum, the mean relative error $\bar{r}$ is 
\begin{equation}  
    \bar{r}_{\bar S, FPE} = 0.025.
\end{equation}
FPE performs well in reconstructing the spectrum even with relatively short time series, with a deviation from the true spectrum around 2\%, but it is evident that it is not able to capture all the features of the spectrum, since the reconstruction in the tails is inaccurate. If we compute the average percentage error in three standard deviations from the mean we gain an order of magnitude estimate of the accuracy of the reconstruction, obtaining: 
\begin{equation}
    \nonumber
    \bar{r}_{\bar S, FPE}^{\mu \pm 3\sigma} = 0.005. 
\end{equation}
and we can state the the method is highly performant in this frequency window. 
Since it is not always possible to work with more then one realization for a process, as in observational experiments, it is important to understand how this method behaves when we apply it on a single time series. In that case, the mean error and its standard deviation are: 
\begin{equation}
    \nonumber
    \bar r_{FPE} = 0.16; \quad \sigma_{r} = 0.16\,.
\end{equation}
The average error of a reconstruction based on a single time series is bigger than the one obtained considering the average of all the reconstructions.
Standard deviation of the errors assume small values and 90\% credibility regions (Cr.) in \ref{fig:FPEmean} are tight. This means that if we randomly take two of the reconstructed spectra, we expect their differences to be statistically small. These facts are evidence for FPE to be a relatively stable optimizer. 
We already said that the estimate for filter's length lays in a well defined range of values. The quality of the reconstruction, as noted in Chapter 1, depends on the order of the reconstructed AR process. 
To investigate this idea further, we plot the value of the optimizer for each recursion and the residuals as a function of the estimate for the order of the autoregressive process (\textbf{Figure \ref{fig:FPEErrorOrder}}),

\begin{figure}[H]
    \noindent
    \centering
    \includegraphics[width = \linewidth]{Images/NormalPSD/FPEoptimizer.pdf}
    \caption{Stability of FPE, that results in a cluster of values in a small region (left), is due to the typical behaviour of the optimizer as a function of the AR order (right).}
    \label{fig:FPEErrorOrder}
\end{figure}
Foremost evidence for the stability of this optimizer is found in its values as a function of the AR order. Figure \ref{fig:FPEErrorOrder} shows that Values tend to decrease monotonically until a sharp and well defined minimum is reached, to start growing again after this area. This is representative of the typical behaviour of this optimizer as a function of the length of the filter, and explains why the estimates for AR orders are confined in a small region. This plot also shows that most of the values collapsed in the region of AR orders comprised between 30 an 50. There are only few values outside this interval, and get fewer as order's estimate increases. In this region, there is no strong evidence of some correlation between the estimate for M and its associated relative error. It is important to notice that the estimate associated with the higher values of M (around $M > 60$) exhibit the highest 'mean' relative error. \\ 
We conclude that the FPE shows good quality reconstruction for the spectrum and very desirable stability properties, its estimate for filter's length M is clustered in a region where there is no dependence of error from M. 
\subsubsection{Optimum Bayes Decision Rule (OBD)}
The second optimizer in consideration is OBD.  (\textbf{Figure \ref{fig:OBD3000}}) shows a good agreement between the average over the reconstructions and true spectrum.
\begin{figure}[H]
    \centering
    \includegraphics[width = \linewidth]{Images/NormalPSD/OBDpsdAndResiduals.pdf}
    \caption{Reconstruction of the spectrum from 1000 simulations with OBD optimizer. Plot is mean of all the results with 90\% credibility regions (above) and percentage error (below)}
    \label{fig:OBD3000}
\end{figure}
The same behaviour of FPE is observed, a good quality reconstruction at the center of the distribution with quite small Credibility regions (Cr.) at a $90$\% credibility level.
Qualitatively they behave similarly but quantitatively the disagreement of this method is found to be even larger than the one obtained with FPE, as can be noticed in the plot of relative errors in figure \ref{fig:OBD3000} (below) where at the extremes the inaccuracy is even larger than 300\% with respect to the real value of the spectrum. For the overall spectrum and within three standard deviation from the mean, we find that the reconstructed spectrum has a deviation of
\begin{equation}
    \nonumber
    r_{\bar S, OBD} = 0.15, \qquad r^{\mu \pm 3\sigma}_{\bar S, OBD} = 0.018. 
\end{equation}
When applied to single realizations of the time series, we find
\begin{equation}
    \nonumber
    \bar{r}_{OBD} = 0.237, \qquad \sigma_{r, OBD} = 0.0296.
\end{equation}
It is interesting to notice that overall relative error of the average spectrum is of the same order of magnitude of the average single reconstruction for the spectrum. Averaging over a 1000 realisations does not substantially improve our results. As a function of the filter length, the values taken by the OBD cluster in a small region, indicating the stability of the reconstructed process order $M$.

>>>>

\usepackage{amsmath,amssymb}
\usepackage{amsfonts}
\usepackage[normalem]{ulem}
\usepackage{textcomp}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{bm}
\usepackage{afterpage}
\usepackage{graphicx}
\usepackage{psfrag}
\usepackage{mathtools}
%\usepackage{tensor}
\usepackage{layouts}
%\usepackage{DejaVuSans}
\usepackage{epstopdf}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[utf8]{inputenc}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{tabularx}
\usepackage{ragged2e}
\usepackage{cleveref}
\usepackage{blindtext}
\usepackage{graphicx}
\usepackage{siunitx}
	\sisetup{output-decimal-marker={.}}
\newtheorem{theorem}{Theorem}
\usepackage{float}

@@

The application of MESA is not limited to spectral estimates, but it also provides a link between spectral analysis and the study
of autoregressive processes.\cite{doi:10.1029/RG013i001p00183}.
An autoregressive stationary process of order $p$ (AR($p$)) is a time series whose values satisfy the following expression: 
\begin{equation} \label{eq:AR_p}
    x_t - b_1 x_{t-1} - b_2 x_{t-2} \dots b_p x_{t - p} = \nu_t
\end{equation}
where $b_1, \ldots, b_p$ are real coefficients and $\nu_t$ is white noise with a given variance $\sigma^2$.
Thus, an AR($p$) process models the dependence of the value of the process at time $t$ on every past $p$ observations, 
thus being potentially able to model complex autocorrelation between data.
\par
To show the analogy, we compute the PSD of such process and we show that it is formally equivalent to the PSD obtained in eq.~\ref{eq:MESApsd}.
We start taking the z \footnote{Spiegare in due righe la z transform} transform of the previous eq.~\ref{eq:AR_p}: 
\begin{align}
    \sum_t x_t z^t - \sum_i b_i z^i\sum_t x_{t - i} z^{t - i} = \sum_t \nu_t z^t.
\end{align}
Calling $\tilde x(z)$ and $\tilde \nu (z)$, the transformed quantities, 
in the $z$ domain, the process takes the form:
\begin{equation}
    \tilde x(z) = \frac{\tilde\nu(z)}{\left(1 - \sum_{i = n}^p b_n z^n \right)}
\end{equation}
Since we assumed a stationary process, $\tilde{x}(z)$ is analytic both on and inside the unit circle. Taking its square value and evaluating it on the unit circle $z = e^{-\imath 2 \pi f \Delta t}$, from the definition of spectral density one obtains:
\begin{equation}\label{eq:ARspectrum}
    S(f) = \vert \tilde x(z)\vert ^ 2 = 
    \frac{\vert \tilde \nu(f) \vert ^ 2}{\left\vert 1 - \sum_{n = 1}^p b_i e^{\imath 2 \pi f n \Delta t} \right\vert ^ 2}.
\end{equation}
The numerator is the spectral density of white noise $\nu_t$, i.e. its (constant) variance $\sigma^2$.
Eq.~(\ref{eq:ARspectrum}) and Eq.~(\ref{eq:MESApsd}) are equivalent, if we identify $b_i = - a_i$ and $P_N \Delta t= \sigma ^ 2$.
This shows that the MAXENT estimation of the PSD computes the coefficients of the autoregressive process that is able to model the data faithfully.

@@

This section is devoted to an extensive study on how to make such choice.
In section~\ref{sec:optimizers}, we are going to define three different \textit{loss functions} to measure how well the algorithm is able to reproduce a known PSD.
The basic idea is to validate, as the autoregressive order considered increase, the performance of the algorithm results by measuring the loss function and pick, among the orders the one that yields better results.
Second, the performance of different loss functions will be assessed by performing the spectral analysis on time series with given analytical PSD, sec.~\ref{sec:arp_validation}.
In a last subsection~\ref{sec:LIGO_validation}, the same analysis is performed on real data from LIGO-Virgo gravitational waves observatory.

\subsection{Choosing an autoregressive order}\label{sec:optimizers}

A good choice for the autoregressive order $M$ is proposed to be \cite{doi:10.1190/1.1440902}
\begin{equation}\label{eq:MMAx}
M_{max} = 2N / \ln{(2N)}
\end{equation}
However, this is just an upper limit for the estimate of the length $M$ and the optimal algorithm could employ much less points.
We then need a more sophisticated method for computing the right value for $M$.
Various loss functions to assess the algorithm performance have been proposed in literature.
We summarize them below:

\begin{itemize}
\item \textbf{Final prediction Error (FPE)} 
The first criterion is due to Akaike (1970) \cite{Akaike1970StatisticalPI}. He proposed that M has to be chosen as the length that reduces the error when the filter is used as a predictor: 
\begin{equation}
    FPE(m) = \mathbb{E}\left[ \left((x_t - \hat x_t) ^ 2\right) \right]
\end{equation}
with $\hat{x}_t = \sum_{i = 1}^M a_i x_{t - i}$.
Minimizing FPE is the same as minimizing the quantity: 
\begin{equation}
    P_{m} \frac{N + m + 1}{N - m - 1}
\end{equation}
with $P_M$ defined by the prediction error filter equation \sschmidt{What is this equation?}.
\item \textbf{Criterion Autoregressive Transfer function (CAT)}
This second loss function is due to Parzen \cite{bhansali1986}, and it is defined as
\begin{equation}
    CAT(m) = \frac{1}{N}\sum_{k = 1}^m \frac{N - k}{N P_k} - \frac{N - m}{N P_m}.
\end{equation}
It has the interesting property that the distance between the original filter and the estimate is minimum and asymptotically unbiased.

\item \textbf{Optimum Bayes Decision rule (OBD)} 
The last criterion we will consider is the OBD (Rao, 1982)\cite{doi:10.1029/WR018i004p01097}. With the technique of asymptotic integration, th loss function is proposed to be the length that maximizes the posterior distribution \sschmidt{Which posterior?}: 
   \begin{align}
        OBD&(m) = (N - m - 2) \log(P_m) \nonumber\\
        &+ m \log(N) + \sum_{k = 0}^{m-1} \log(P_k) + \sum_{k = 1}^{m} a_k^2.
    \end{align}
\end{itemize}

Once a loss function is selected, the choice of the best recursion order is straightforward: we solve the Levinson recursion until $M_{max}$, as given in Eq.~\ref{eq:MMAx}, iterations are reached. Then, the order $M$ is selected to be the one that minimize the loss function.

In a real implementation of the algorithm, computing all the recursion up to $M_{max}$ can result in a significant waste of computational power: the optimal value is often $M_{opt}<< M_{max}$ and, in such cases, computing all the values of $m$ until $M_{max}$ is not useful.
In practice, we can apply an \textit{early stopping} procedure: every few iterations we look for the best order of $M_{opt}$; if this value does not change for a while, the computation is stopped and a good (local) minimum of the loss function is found.

The following sections will be devoted to the study of the statistical properties of the loss functions introduced above: we need to understand which choice provides the best quality in the reproduction of some known power spectral densities.

\subsection{Choosing a loss function - A gaussian PSD} \label{sec:arp_validation}
We test the performance of the three loss functions on a random time series with a known power spectral density.
We take the PSD to be a Gaussian distribution with mean $\mu = 2.5 \quad units$ and standard deviation $\sigma = 0.5 \quad units$, see fig.~\ref{fig:autocorr}.
Of course, the time series will be generated by sampling a frequency vector from ${p(\tilde{n}(f_i)) \propto \exp{\left\{-\frac{1}{2}\sum_i\frac{n_i ^2}{S(f_i)}\right\}}}$ and performing the inverse Fast Fourier Transform on the frequency series.

We will apply Burg method with each loss function on a number of simulated time series, obtaining a variety of estimate for PSD.
The disagreement between the PSD $S_i(f)$ estimated from the i-th simulated time series and the target PSD $S(f)$ is measured with the relative error $r_i$, averaged over all frequencies:
\begin{equation}
r_i = \frac{1}{f_{Ny}}\int_0^{f_{Ny}} \d{f} \frac{|S_i(f) - S(f)|}{S(f)}
\end{equation}
where $f_{Ny}$ is the Nyquist frequency of the time series. Of course in practice, the integral will be computed by averaging over a discrete grid of frequencies.

We will consider every loss function separately and compare the results only at the end of this section.
For each loss function, we will compute the ensemble averaged PSD estimation (as well as the $90\%$ confidence level) and the frequency dependence of the error (ensemble averaged):
\begin{equation}\label{eq:ensemble_averaged_error}
	r(f) = \sum_i \frac{|S_i(f) - S(f)|}{S(f)} \;.
\end{equation}
In figure~\ref{fig:FPEmean} \ref{fig:OBDmean} and \ref{fig:CATmean} we report the averaged PSD for the three loss functions, as well as the ensemble averaged error in eq.~\ref{eq:ensemble_averaged_error}.

Furthermore, we display in a contour plot the correlation between $r_i$ and the autoregressive order $m_i$ chosen for each independently drawn time series.
This is done in figure~\ref{fig:optcomparison}.

For our studies, we generate a dataset of 1000 time series of 3000 points each: see fig.~\ref{fig:noise} for an example of a time series in frequency and time domain.
In section~\ref{sec:LIGO_validation}, we will perform the analysis again using the Advanced LIGO design sensitivity PSD \cite{Ligo}.

@@

\begin{figure}
    \centering
        \includegraphics[width = \linewidth]{Images/NormalPSD/NormalNoise.pdf}
    \caption{Noise in both time and frequency domain. Frequency noise behaviour is representative of the chosen spectrum, having its maximum amplitude around $f = 2.5 au$ \sschmidt{Togliamo questa figura?}}
    \label{fig:noise}
\end{figure}

\begin{figure}[t]
	\centering
	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/FPE_spectrum_estim.pdf}
	\caption{Ensemble Average for FPE reconstruction with 90\% Credibility Regions (above) and percentage error (below)}
	\label{fig:FPEmean}
\end{figure}
\begin{figure}[t]
	\centering
	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/OBD_spectrum_estim.pdf}
	\caption{Reconstruction of the spectrum from 1000 simulations with OBD optimizer. Plot is mean of all the results with 90\% credibility regions (above) and percentage error (below)}
	\label{fig:OBDmean}
\end{figure}

\begin{figure}[t]
	\centering
	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/CAT_spectrum_estim.pdf}
	\caption{Mean spectrum for CAT optimizer, 1000 simulations for 3000 points datasets}
	\label{fig:CATmean}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width = \linewidth]{Images/optimisers_comparison/normal/error_length_contour.pdf}
	\caption{Relative error as function of the length of the filter.}
	\label{fig:optcomparison}
\end{figure}

\paragraph{Final Prediction Error (FPE)}
The results of our investigation on the performance of FPE are reported in fig.~\ref{fig:FPEmean}.
Qualitatively, there is a good agreement between the reconstructed spectrum and the true spectrum; however, we note that the reconstruction is not very accurate in the low frequency region. 
Furthermore, the $90\%$ credible region is very small: this means that if we randomly take two of the reconstructed spectra, we expect their differences to be statistically small.
These facts are evidence for FPE to be a reliable loss function.

By looking at the red series in fig.~\ref{fig:optcomparison}, we note that the autoregressive orders obtained with FPE are clustered in a very small region. This is a desiderable property: FPE provides a stable estimation of the autoregressive order, which does not affect much the reconstruction error.
We conclude that the FPE shows good quality reconstruction for the spectrum and very desirable stability properties, its estimate for filter's length M is clustered in a region where there is no dependence of error from M. 
\paragraph{Optimum Bayes Decision Rule (OBD)}
The second loss function in consideration is OBD.
The results in fig.~\ref{fig:OBDmean} show a good agreement between the average over the reconstructions and the true spectrum.
Qualitatively the same behaviour of FPE is observed: a good quality reconstruction at the intermediate and high frequencies with a low $90\%$ confidence level; and a worse performance at low frequencies.
However, when looking at the error, the disagreement of this method is found to be larger than the one obtained with FPE: in the worst case, the error can gain a factor of $2$ as compared with FPE.

When plotting the error as a function of the autoregressive length (green series in fig.~\ref{fig:optcomparison}), the values taken by the OBD cluster in a small region, indicating the stability of the reconstructed process order $M$. We note that on average, OBD tends to choose smaller values of $M$ with respect to FPE: this can probably explain the observed large errors.

\paragraph{Criterion Autoregressive Transfer function (CAT)}
The performance of CAT are shown in fig.~\ref{fig:CATmean} and it is striking that CAT differs substantially from the other two loss functions considered.
The ensambled average PSD matches very well the ``true" PSD. This is also true even in the low frequency region, where the OBD and FPE had a poor performance.
However, the variance of the reconstructed spectrum is quite large (much larger than for FPE and OBD), and the relative error is quite high $~10\%$ and it approximately constant over all the frequencies.

The reason for this behaviour, can be found in fig.~\ref{fig:optcomparison} (black series). CAT does not converge to any specific value for the order of the autoregressive process.
In turn, the estimates for the order span a large range of values, hence the large variance observed.
Figure~\ref{fig:optcomparison} also shows a strong dependence of the error on the estimated length of the filter. The good quality of the reconstruction from the average spectrum can be explained as follows: long filters are able to capture features that short filters cannot see, like outliers in different realization of a noisy timeseries, but this is also responsible of an augmented variance in the estimate, by introducing spurious peaks in the reproduction.

When averaging the diffrerent PSD estimation, the noise in each spectrum cancels. It is interesting that the cancelling of errors in averaging is property of gaussian noise (the process is gaussian, after all). This implies, in a sense, that each estimate of the spectrum is independent of any other, as suggested by the huge variance in the residuals. This instability is not a good property for an estimation, so that this estimator might be found to be optimal for repeatable experiments and not for single observation of some process.

\paragraph{Final remarks on the choice of the loss function}

In our analysis, FPE and OBD loss functions are found to behave similarly while CAT shows completely different properties.
CAT provides an accurate average spectrum over all the frequencies and large amount of variability; in turn OBD and FPE provide a poorer average in the low frequency tails with a lower variability (FPE with less noise than OBD).

The bad behavior in the reconstruction of the tails from OBD and FPE might be due to the fact that very small values require longer filters. 
In fact when we consider the average for all the spectra we note that short filters yield larger the error in the tails. OBD, which select the shortest filters, can provide an error as large as $300 \%$ at the extrema, as compared with $85\%$ error od FPE. In turn, CAT is $5\%$ accurate in those regions.

However, we see in figure \ref{fig:optcomparison} that a single PSD estimation with FPE is still able to provide the lowest averaged error over all frequencies, while CAT can reach error error $5$ times larger.
This is to say that CAT is the loss function that minimizes errors in the tails, FPE is found to reach best overall accuracy.

The conclusion is clear: even if in some cases CAT might be more accurate when taking the average over several estimation, FPE guarantees that the single estimation will be more accurate.
As in any common situation we cannot perform such averaging over different realization of the same time series, we must prefer FPE over CAT (let alone OBD, which even though qualitatively similar to FPE has worse performance).
However, if we indeed can measure the PSD by averaging over different time series, probably using CAT as a loss function is the best choice.
\sschmidt{This is similar to Welch: averaging over different time series to get a reliable result. Should we mention it?}
\sschmidt{C'erano ancora molti commenti: ho sintetizzato molto. Va bene? O si vuole aggiungere altro?}

%It is evident that tails do not affect CAT  average error and standard deviation at all, so that it's accuracy is approximately constant everywhere in the spectrum. \\ \ 
%Since both FPE and CAT shows pros and cons, at this level, there is no reason to choose between them.\\ 
%Analyzing the resolution of the single reconstruction, figure \ref{fig:optcomparison} clearly shows that FPE minimize overall relative error giving a stable result.\\ 
%CAT instead shows a larger error and do not converge to any specific value. Almost every reproduction obtained with CAT is worst with respect to those obtained with FPE, but their mean is more precise. \\ 
%It's instability is probably the cause for this behaviour. Since it does not show a strong convergence to some order with respect to others, very different lengths are selected.
%Longer lengths are able to capture a lot of details but are more likely to introduce noise. In other words, long filters reduce bias but increment the variance of the result. 
%Taking the average, for central limit theorem the error decreases as $\sim \frac{1}{\sqrt{N}}$. In this way we compensate noise and obtain a spectrum with small bias and small fluctuations.
%Considering the average, CAT reconstruct precisely and with good accuracy every part of the spectrum, even if in "regular" parts best accuracy is reached from FPE. For a single reproduction, CAT is very noisy and unstable, while FPE performs the better. \ref{fig:optcomparison}
%definitely shows that FPE gives the best results: it is a stable method that choose filters in the clustered area associated with minimum error, while both OBD and CAT result lies outside this area and show bigger errors. Two other interesting plots that stress this fact are reported in figure \ref{fig:ordersCompairson} and \ref{fig:residualsComparison}
%where the histograms for the errors and for order's estimate are reported. These graphs provide stronger evidence for our previous statements. As already noticed, FPE is very stable and provide the smallest 'overall relative error'. \\
%CAT histogram shows a very interesting behaviour. The histogram for the order show two different peaks,  a small one associated to 'short filters', and a second high peak at $m = M$. This means that CAT is most likely to not converge at all at some value for the regression order, and it is more likely selects a length that is comparable with the largest available.  If short filters are selected, the overall percentage error will be low, respectively represented by the first peak in the distribution of the errors, while if long filters are selected, the error is maximized.

\subsection{Choosing a loss function - LIGO Spectrum} \label{sec:LIGO_validation}
\begin{figure}
    \centering
    \includegraphics[width = \linewidth]{Images/LIGOsimulate/LigoPSDoriginal.pdf}
    \caption{Original power spectral density from Ligo}
    \label{fig:LigoPSDOriginal}
\end{figure}
In this subsection we concentrate again on the reconstruction of a specific, known, power spectrum.
We analyze the properties of reconstruction of this second spectrum whose shape is the Advanced LIGO design sensitivity theoretical spectral curve (Figure \ref{fig:LigoPSDOriginal}). 
For this analysis we will study properties only for arrays of 40960 points. We consider a sampling rate of $2048 samples / second$ that means we can reconstruct a Nyquist frequency of $1024 Hz$. We chose a longer dataset than before because algorithm works with a 1 dimensional interpolation of spectrum, and with shorter datasets the interpolated spectrum is not consistent with the original spectrum and does not capture the height of the peaks. Shorter choices would result in noise extracted from a different power spectrum than the one in figure \ref{fig:LigoPSDOriginal}. \\ 
We will study statistical properties from samples obtained with 500 simulations and compare the obtained results for each optimizer. 
We start considering statistical properties of mean spectrum obtained averaging over all the simulations and conclude with an analysis for the statistics of the reproduction of single spectrum. 
\subsubsection{Properties for the average spectrum}
Here we report mean properties of the reproductions of the spectrum. The spectrum we want to reconstruct is reported in Figure \ref{fig:ligospectrum},
\begin{figure}
    \centering
